---
title: "Stock Market Prediction Project - Advanced Dataset"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---

```{r}
library(ISLR)
library(caTools)
library(MASS)
library(class)
library(forecast)
library(ggplot2)
library(TTR)
library(fpp2)
library(zoo)
```

```{r}
smp=read.csv("/Users/bonnie/Desktop/Smarketadvanced.csv", header=T, sep=',', 
         na.strings="NA")
attach(smp)
dim(smp)
```

Our goal is to create insights and draw predictions for the percentage returns from the advanced stock market dataset over 801 months (66 years and 9 months) in the past, from January 1952 to September 2018. 

For each month, we have recorded: 

1. Direction: whether the market was Up or Down of this month 

2. MktReturn: market percentage return of the month 

3. X12m_cpi_forecast: CPI Forecast on a semi-annual basis

4. X12m_real_gdp_forecast_calculated: calculated real GDP forecast on a semi-annual basis 

5. X12m_unemployment_rate: unemployment rate on a semi-annual basis

6. wti_prev_yoy_chg: WTI previous year-over-year change 

7. PE_ratio_shiller: Shiller PE ratio for the S&P 500 

8. corp_bond_yield_prev_month: corporate bond yield in previous month 

9. MktReturn_prev_month: market return of the previous month from the date

10. MktReturn_prev_yr: market return of the previous year from the date

*Slide Notes on Financial Terminology:*

* CPI (Customer Price Index): a measure that examines the weighted average of prices of a basket of consumer goods and services, such as transportation, food, and medical care)

* Shiller PE ratio: a market valuation indicator, usually for S&P 500

* WTI: Crude Oil Price 


```{r}
summary(smp)
```


Next, using a naive strategy, we can produce a matrix that contains all of the correlations between market return and all other predictors in the data set. 

```{r}
cor.mat=cor(smp[, -2])
cor.mat[,1:2]
```

As we can see, the correlations between all the predictor variables and market montly returns are close to zero. Only the variables X12m_real_gdp_forecast_calculated, X12m_unemployment_rate, wti_prev_yoy_chg, PE_ratio_shiller, MktReturn_prev_month have correlations that is relatively larger and close to 0.1. 

##### Logistic Regression
Next, we want to predict Direction by fitting a logistic regression model using the predictor variables that have a relatively high correlation with the market return. 

```{r}
glm_fit=glm(Direction~X12m_real_gdp_forecast_calculated+X12m_unemployment_rate+wti_prev_yoy_chg+PE_ratio_shiller+MktReturn_prev_month, data=smp, family=binomial)
summary(glm_fit)
```
Assume the threshold for small p-value is 0.05. Then, only the p-value of X12m_real_gdp_forecast_calculated suggests that there is a strong evidence showing there are some relationship between Direction and X12m_real_gdp_forecast_calculated. 

Then, we can improve our model by reducing insignificant variables. 

```{r}
glm_fit=glm(Direction~ X12m_real_gdp_forecast_calculated, data=smp, family=binomial)
summary(glm_fit)
```

```{r}
glm.probs=predict(glm_fit,type='response')
contrasts(Direction)
glm.probs[1:10]
max(glm.probs)
min(glm.probs)
```
We then can get the predicted probability of the market going up from the logistic regression model. The range is between 48.22% and 75.76%.


Next, we want to convert the above predicted probability to class label in order to make a prediction for the market direction going up or down on a particular day. So, we create a vector of class predictions based on whether the predicted probability of a market increase is greater than or less than 0.5. 

```{r}
glm.pred=rep("Down",801)
glm.pred[glm.probs >.5]="Up"
table(glm.pred,Direction) #confusion matrix
```
```{r}
mean(glm.pred==Direction )
```
This logistic regression model correctly predicted the movement of the market 63.05 % of the time. However, the training error rate is 36.95%, which might be overly optimistic. 

To better access the accuracy in this model, we split the data into a training set and a test set holding 30% of data for testing. 
```{r}
set.seed(101)
sample = sample.split(1:nrow(smp), SplitRatio=0.3) 
test = subset(smp, sample==TRUE)
train = subset(smp, sample==FALSE)
direction.test=test$Direction
glm_fit=glm(Direction~X12m_real_gdp_forecast_calculated, data=train, family=binomial)
summary(glm_fit)
```
```{r}
glm.probs=predict(glm_fit,test, type='response')
glm.pred=rep("Down",240)
glm.pred[glm.probs >.5]="Up"
table(glm.pred,direction.test)
mean(glm.pred==direction.test)
mean(glm.pred!=direction.test)
152/(152+79)
```
In this improved model, we got a (slightly) better prediction rate of 63.75%, and we lowered the test set error rate to 36.25%. 

The confusion matrix also shows that on days when logistic regression predicts an increase in the market, it hasa 65.8% accuracy rate. This suggests a possible trading strategy of buying on days when the model predict an up market, and avoid trades on days when a down is predicted. 


##### LDA 
```{r}
lda.fit=lda(Direction~X12m_real_gdp_forecast_calculated, data=train)
lda.fit
plot(lda.fit) #Plot the linear discriminants for each training observation
```
The LDA output indicates that 38.3% of the training observations correspond to the months during which the market went down. 

LDA decision rule: $\delta_k(x)=x^T\sum^{-1}\mu_k - \frac{1}{2}\mu_k^T\sum^{-1}\mu_k + log(\pi_k)$

If 0.512 * X12m_real_gdp_forecast_calculated is large, then the LDA classifier will predict a market increase; and if it's small, then the LDA classifer will predict a market decline. 

```{r}
lda.pred=predict(lda.fit, test)
lda.class=lda.pred$class
table(lda.class, direction.test)
mean(lda.class==direction.test)
```
The LDA and logistic regression predictions are almost the same. 

Then applying a 50% threshold to the posterior probabilities to recreate the predictions contained in lda.class. 

```{r}
lda.pos=lda.pred$posterior
sum(lda.pos[,1]>=.5)
sum(lda.pos[,1]<.5)
lda.pos[1:20,1] #posterior probability of a decreasing market 
lda.class[1:20]
sum(lda.pos[,1]>=.5651)
```
The greatest posterior probability of decrease in the test data was 56.51%. 
???

##### QDA

```{r}
qda.fit=qda(Direction~X12m_real_gdp_forecast_calculated, data=train)
qda.fit
```
```{r}
qda.class=predict(qda.fit,test)$class
table(qda.class, direction.test)
mean(qda.class==direction.test)
```
The QDA predictions are accurate 66.25% of the time, which has been improved a lot from all the other models. This suggests that the quadratic form assumed by QDA may be more accurate than the linear form assumed by LDA and logistic regression. 


##### KNN
```{r}
train.gdp=cbind(train$X12m_real_gdp_forecast_calculated)
test.gdp=cbind(test$X12m_real_gdp_forecast_calculated)
train.dir=train$Direction
```


```{r}
test.error.rate <- c()
k_value=c(1,5,10,15,20,30,50,100,150,200) 
for (i in k_value){
knn_model=knn(train=train.gdp, test=test.gdp, cl= train.dir, k=i)
test.error.rate=c(test.error.rate, mean(knn_model != direction.test)) 
}
test.error.rate
```
KNN performs best around K=5, which gives a 42.08% accuracy rate. 


##### Time Series

```{r}
market_return=smp[,3]
mr_ts=ts(market_return,frequency=12, start=c(1952,1))
plot.ts(mr_ts)
```
Varing spread - the time series does not have constant variance. 
Fail stationary criterion. 

```{r}
plot(aggregate(mr_ts,FUN=mean), ylab="yearly_market_return")
# a year on year trend
```


```{r}
bp=boxplot(mr_ts~cycle(mr_ts)) #seasonal effect
```
Inferences: 
1. The year on year trend shows that there is a periodical change in the stock market index.
2. The mean value in November and December is higher than the rest of the months. The variance in October is much higher than the rest. 
3. The mean value in each month is quite similar and their variance is small. Hence, we have a seasonal effect with a cycle of 12 months or less. 

```{r}
ts_lag1=lag(mr_ts, lag=1)
head(cbind(mr_ts, ts_lag1))
ts_diff=diff(mr_ts, lag=1)
tm=cbind(mr_ts, ts_diff)
head(tm)
plot.ts(tm) 
```

```{r}
library(tseries)
adf.test(ts_diff) 
```
By taking the difference, we make the non-stationary series stationary. 


Plot ACF/PACF charts
```{r}
acf(ts_diff)
```


```{r}
pacf(ts_diff)
```

```{r}
autoArimaFit=auto.arima(ts_diff)
plot(forecast(autoArimaFit, h=20))
```


```{r}
arimaFit=Arima(ts_diff,order=c(3,1,0))
plot(forecast(arimaFit,h=20))
```

```{r}
tbatsFit=tbats(ts_diff, use.parallel=T, num.cores=2) # fit tbats model
plot(forecast(tbatsFit)) # plot
components <- tbats.components(tbatsFit)
plot(components)
```

Confidence Interval for my Forecasts
```{r}
model=HoltWinters(ts_diff) 
predict(model, 50, prediction.interval=T, level= 0.99)
```





